---
title: "Fixed Effects Estimation"
author: "Balachandar & Mohammed"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning=F)
```

## Panel Data Analysis with R
## Instructions:

- Please replace --- with appropriate function or variable name to run chunk without any error

## 1. Load the library

```{r}
library(plm)
library(knitr)
library(broom)
library(tidyverse)
library(stargazer)
library(lmtest)
library(gplots)
library(foreign)
library(strucchange)
```


## 2. Import Dataset (STATA version)

# Rental Data from Wooldridge

```{r}
RENTAL <- read.dta("RENTAL.DTA")

# Looking at the structure of the data
str(RENTAL)
```

# Information about the data set:

city: city label, 1 to 64
year: 80 or 90
pop: city population
enroll: # college students enrolled
rent: average rent
rnthsg: renter occupied units
tothsg: occupied housing units
avginc: per capita income
lenroll: log(enroll)
lpop: log(pop)
lrent: log(rent)
ltothsg: log(tothsg)
lrnthsg: log(rnthsg)
lavginc: log(avginc)
clenroll: change in lrent from 80 to 90
clpop: change in lpop
clrent: change in lrent
cltothsg: change in ltothsg
clrnthsg: change in lrnthsg
clavginc: change in lavginc
pctstu: percent of population students
cpctstu: change in pctstu
y90: =1 if year == 90

#  Indexed by city and year

```{r}
# Converting the data frame to panel data frame to work with plm package:
# After converting the data frame into a panel data please check the structure of the data frame

rental_p <- pdata.frame(RENTAL, index=c("city", "---"))
```

## 3. Display the data

```{r}
# Looking at the structure of RENTAL and p_rental data frame:
str(RENTAL)
str(rental_p)

# Check the first few rows of the data frame:
head(RENTAL)
head(rental_p)
```

#  Model lrent~y90+lpop+lavginc+pctstu

## 4. OLS using lm

```{r}
ols <- lm(lrent~y90+lpop+lavginc+pctstu, RENTAL)

summary(ols)
```

## 5. OLS using plm

```{r}
# This format uses the panel data frame we actually loaded 
# When we use the data frame without panel data conversion - we need to specify the random effects variables as index:
# What is missing here? 

pooled <- plm(lrent~y90+lpop+lavginc+pctstu, data=RENTAL, model="pooling", index=c("city", "year"))
summary(pooled)

#OR use this format
# This format uses the panel data frame we created using plm package
# Run Pooled OLS

pooled2 <- plm(lrent~y90+lpop+lavginc+pctstu, data=rental_p, model="----")

summary(pooled2)
```

## Results table
```{r}
# Comparing the results of both approaches:

stargazer(pooled, pooled2, type ="text")
```

## 6. Test for heteroscedasticity

```{r}
res <- residuals(ols)
yhat <- fitted(ols)

plot(RENTAL$pctstu, res, xlab = "%Students", ylab = "Residuals")
plot(yhat, res,  xlab = "%Students", ylab = "Residuals" )
```

## 7. Fixed Effects

# Includes within-entity effects
```{r}
# specifying model="within" runs the model by demeaning the data within each group

fe <- plm(lrent~y90+lpop+lavginc+pctstu, data=rental_p, model="within")

summary(fe)

# Show fixed effects for all 64 cities
# fixef(fe) # run this code if you wish see the fixed effects estimates of each cities - just remove the # in front of fixef(fe) to run the code
````

## 9. Test for FE vs OLS

# Ho: OLS is better than FE, reject at p < 0.05
```{r}

pFtest(fe, ols)
stargazer(pooled, fe, type = "text")

# Why estimates for constant is available for the fixed effects estimator model?
```

## 10. Chow test for FE vs OLS

# Ho: OLS is better than FE, reject at p < 0.05

```{r}
sctest(RENTAL$lrent ~ RENTAL$pctstu, type = "Chow", point = 10)

# infer the results
```

# ------------------------------- DEMO 1 ENDS HERE -----------------------------------#

## 9. Random Effects

# Includes both the within-entity and between-entity effects

```{r}
# by specifying model="random", we run the random effects model:

re <- plm(lrent~y90+lpop+lavginc+pctstu, data=rental_p, model="random")

# Look at the summary of the results of the random effects model

---(re)
```

## 10. FE VS RE

## Hausman Test Ho: RE is preferred, Ha: FE is preferred (p < 0.05)

```{r}
phtest(fe, re)
```

## 12. Breusch Pagan Lagrange Multiplier Test Ho: No panel effect, i.e., OLS is better. Ha: RE is better at p <0.05

```{r}
plmtest(pooled, type = c("bp"))

#plmtest(ols, type=c("bp")) 
```

## Example 2: Random effects estimator model:

## Loading the data set "Grunfeld" from plm package:

```{r}
data("Grunfeld", package="plm")
```

## Data description:

1. firm - observation
2. year - date
3. inv - gross Investment
4. value - value of the firm
5. capital - stock of plant and equipment

## Looking at the first few rows of the data frame:

```{r}
head(Grunfeld)
```

## Converting the data frame into a panel data frame suitable to working with plm package:

```{r}
p_grunfeld <- pdata.frame(Grunfeld, index=c("firm","year"), drop.index=TRUE, row.names=TRUE)

head(p_grunfeld)
```

## Run fixed effects model and random effects model:

```{r}
grun.fe <- plm(inv ~ value+capital, data = Grunfeld, model = "within")
grun.re <- plm(inv ~ value+capital, data = Grunfeld, model = "random")
```

## Print the results of random effects model:

```{r}
summary(grun.re) # right now the model does not explain the time error component - let's do it in the next chunk
```

## Fixed effects model vs. Random effects model - Hausman Taylor test:

## Hausman Test Ho: RE is preferred, Ha: FE is preferred (p < 0.05)

```{r}
phtest(grun.fe, grun.re)

```

# Advanced Panel effects estimation:

# Run random effects model with two-ways effects:

```{r}
# This estimation of the twoways error component is based on T. Amemiya (1971)
# The default method is "swar" from Swamy and Arora (1972) - however, it does not estimate the error for time
# Specify the DV to complete the code

grun.tways <- plm(--- ~ value+capital, data = Grunfeld, effect = "twoways",
                  model = "random", random.method = "amemiya")

summary(grun.tways)
```

# Other diagnostic tests: 


## 13. Test for cross-sectional dependence 
# [NOTE: Suitable only for macro panels with long time series] [Not suitable for RENTAL dataset]
# Breusch-Pagan LM test of independence and Pasaran CD test
# Ho: There is no cross-sectional dependence

```{r}
pcdtest(grun.fe, test = c("lm")) # Breusch-Pagan LM test

pcdtest(grun.fe, test = c("cd")) # Pasaran CD test
```


## 14. Testing for serial correlation 
# [NOTE: Suitable only for macro panels with long time series] [Not suitable for RENTAL dataset]
# Ho: There is no serial correlation

```{r}
pbgtest(grun.fe)
```

## 15. Breusch - Pagan test for heteroscedasticity 
# Ho: Homoscedasticity 
# Ha: Heteroscedasticity

```{r}  
bptest(inv~value+capital, data = Grunfeld, studentize = F)
```


